---
title: 'mkvsubs'
author: 'fnaufel'
email: 'https://fnaufel.github.io/'
date: '   (v. `r format(Sys.Date(), "%d/%m/%Y")`)'
lang: 'en'

output: 
  # To install these output formats, run
  #   install.packages("devtools")
  #   devtools::install_github("fnaufel/fnaufelRmd")
  fnaufelRmd::html_report:
    []
---

```{r setup, include=FALSE}
# The next command configures MANY things and loads quite a few packages.
# 
# If you want to see what's being done, execute 
# 
#   cat(
#     system.file(
#       "rmarkdown/resources/R/_common_report.R", 
#       package = "fnaufelRmd"
#     )
#   )
# 
# to find out the location of the file. Then open the file.
# 
# If you want to change the configuration, copy the file, edit it, and
# source it instead of the package file. 
# 
# Or simply write your commands here in this code chunk.

source(
  system.file(
    "rmarkdown/resources/R/_common_report.R",
    package = "fnaufelRmd"
  )
)
```


# Goal

* Extract subtitles from an mkv file.


# Usage

```
mkvsubs.R LANG MKVFILE
```

# Details

* If the track with the specified language does not exist, an error message is given and execution ends.

* If a single track with the specified language exists, a file will be created in the same directory and with the same name as the mkv file --- but with the srt extension.

* If more than one track with the specified language exists, the user will be presented those tracks and their names, from which they must choose one.
  
* If a previous srt file exists, the user will be asked to confirm that the file should be overwritten.


# Tasks

## Call mkvinfo

```{r}
f <- 'Prehistoric.Planet.2022.S01E01.720p.WEB.x265-MiNX.mkv'

info <- system2('mkvinfo', f, stdout = TRUE)

info
```


## Find track and name

### Using grep on the entire output

```{r}
l <- 'por'

matches <- grep(
  paste0('+ Language: ', l),
  info
)

matches
```


### Filter only number, name, language, type (USE THIS ONE)

* `+ Track number: ... (track ID for mkvmerge & mkvextract: ...)`

* `+ Name: ...`

* `+ Language: ...`

* `+ Track type: subtitles`

```{r}
pattern <- paste0(
  '\\+ (',
  '(Track number: )|',
  '(Name: )|',
  '(Language: )|',
  '(Track type: )',
  ').*$'
)

matches <- grep(
  pattern,
  info
)

matches

info <- info[matches]

info
```


## Build data frame

* Process each block of $4$ consecutive lines.

* Get track number.

* Get name.

* Get language.

* Get type.

```{r}
track_no <- vector('numeric')
track_name <- vector('character')
track_lang <- vector('character')
track_type <- vector('character')

i <- 1
j <- 0

while (i <= length(info)) {
  
  # Look for track number
  found_trackno <- regexpr(
    '\\+ Track number: .*: ([[:digit:]]+)\\)$',
    info[i],
    perl = TRUE
  )
  
  # If not found, try next line
  if (found_trackno == -1) {
    i <- i + 1
    next
  }
  
  # Found track number: save
  j <- j + 1
  start <- attr(found_trackno, 'capture.start')[1]
  stop = start + attr(found_trackno, 'capture.length')[1] - 1
  track_no[j] <- substr(info[i], start, stop)

  # Must find name
  i <- i + 1
  found_name <- regexpr(
    '\\+ Name: (.+)',
    info[i],
    perl = TRUE
  )
  
  # If not found, error
  if (found_name == -1) {
    stop('Name not found for track ', track_no[j])
  }
  
  # Found name: save
  start <- attr(found_name, 'capture.start')[1]
  stop = start + attr(found_name, 'capture.length')[1] - 1
  track_name[j] <- substr(info[i], start, stop)

  # Must find language
  i <- i + 1
  found_lang <- regexpr(
    '\\+ Language: (.+)',
    info[i],
    perl = TRUE
  )
  
  # If not found, error
  if (found_lang == -1) {
    stop('Language not found for track ', track_no[j])
  }
  
  # Found language: save
  start <- attr(found_lang, 'capture.start')[1]
  stop = start + attr(found_lang, 'capture.length')[1] - 1
  track_lang[j] <- substr(info[i], start, stop)
  
  # Must find type
  i <- i + 1
  found_type <- regexpr(
    '\\+ Track type: (.+)',
    info[i],
    perl = TRUE
  )
  
  # If not found, error
  if (found_type == -1) {
    stop('Type not found for track ', track_no[j])
  }
  
  # Found type: save
  start <- attr(found_type, 'capture.start')[1]
  stop = start + attr(found_type, 'capture.length')[1] - 1
  track_type[j] <- substr(info[i], start, stop)
  
  # Next block
  i <- i + 1
  
}

attributes(track_no) <- NULL
attributes(track_name) <- NULL
attributes(track_lang) <- NULL
attributes(track_type) <- NULL

df <- data.frame(
  track_no,
  track_name,
  track_lang,
  track_type
)
```

```{r}
df
```

## Filter dataframe

```{r}
found <- subset(df, track_lang == l & track_type == 'subtitles')
```

```{r}
found
```


## If no matches, error

```{r}
if (nrow(found) == 0)
  stop('No subtitle tracks for "', l, '" language found.')
```


## If more than one match, user chooses

```{r}
# Choose the first (hopefully the only) track
choice <-  1
```

```{r eval=FALSE}
# If more than one, show menu
if (nrow(found) > 1) {
  
  choice <- menu(
    found$track_name, title = 'Several tracks found.\nChoose one:'
  )
  
  if (choice == 0) {
    stop('No track selected.')
  }
  
}
```


## Assemble and run command

```{r}
srt_name <- paste0(
  substr(f, start = 1, stop = nchar(f) - 4),
  '.srt'
)
```

```{r}
track_spec <- paste0(
  found[choice, 'track_no'],
  ':',
  srt_name
)
```

```{r}
if (file.exists(srt_name)) {
  
  overwrite <- menu(
    choices = c('Yes', 'No'),
    title = paste0(
      'File exists:\n    ',
      srt_name, '\n',
      'Overwrite?'
    )
  )
  
}
```

```{r}
system2(
  'mkvextract',
  c(
    f,
    'tracks',
    track_spec
  )
)
```

